### 2018-01-10 submission/manuscript.Rmd
update methods section to include top 5 component in workflow

***Study Analysis Overview:***
$\alpha$-diversity was first assessed for differences between controls, adenoma, and carcinoma using both linear mixed-effect models and RR. $\beta$-diversity was then assessed for each individual study for differences between control-adenoma and control-carcinoma. Next, four specific CRC-associated genera (*Fusobacterium*, *Parvimonas*, *Peptostreptococcus*, and *Porphyromonas*) were assessed for differences in RR for adenoma and carcinoma. We then built Random Forest models based on all or CRC-associated genera and these models were trained on one study then tested on the remaining studies, for every study. A similar approach was then applied at the OTU level with the exception that a 10-fold CV over 100 different models, based on random 80/20 splitting of the data, was used to generate a range of expected AUCs. For these OTU based models, the CRC-associated genera included all OTUs that had a taxonomic classification to *Fusobacterium*, *Parvimonas*, *Peptostreptococcus*, or *Porphyromonas*. Finally, the power of each study was assessed for an effect size ranging from 1% to 30% and an estimated sample size, for these effect sizes, was generated based on 80% power. For comparisons in which normal versus adenoma were made the carcinoma samples were excluded from each respective study. Similarly, for comparisons in which normal versus carcinoma were made the adenoma samples were excluded from each respective study. The data was split between feces and tissue samples. Within the tissue groups the data was further divided between samples from the same individual (matched) and those from different individuals (unmatched). Where applicable for each study, predictions for adenoma and carcinoma were then tested for feces, matched tissue, and unmatched tissue.

***Obtaining CRC-associated Genera:***
For the CRC-associated genera analysis of the RR, the total average counts, for 100 different subsamplings, were collected for each respective OTU that had a genus level taxonomic classification to *Fusobacterium*, *Parvimonas*, *Peptostreptococcus*, and *Porphyromonas*. The OTU based Random Forest Models using CRC-associated genera utilized a similar approach except that the OTUs were not aggregated together by genus but kept as separate OTUs. OTU Random Forest models using the full community included all OTUs while those using CRC-associated genera included only those OTUs that had a genus level taxonomic classification to *Fusobacterium*, *Parvimonas*, *Peptostreptococcus*, and *Porphyromonas*.

>>>>

***Study Analysis Overview:***
$\alpha$-diversity was first assessed for differences between controls, adenoma, and carcinoma using both linear mixed-effect models and RR. $\beta$-diversity was then assessed for each individual study for differences between control-adenoma and control-carcinoma. Next, all common genera were assessed for differences in RR for adenoma and carcinoma. We then built Random Forest models based on all or a selected model, based on the top 5 increased and top 5 decreased RR based on P-value, and these models were trained on one study then tested on the remaining studies, for every study. A similar approach was then applied at the OTU level with the exception that a 10-fold CV over 100 different models, based on random 80/20 splitting of the data, was used to generate a range of expected AUCs. For these OTU-based models, the selected model included all OTUs that had a taxonomic classification to a variable in the top 5 increased and top 5 decreased RR based on P-value. Finally, the power of each study was assessed for an effect size ranging from 1% to 30% and an estimated sample size, for these effect sizes, was generated based on 80% power. For comparisons in which normal versus adenoma were made the carcinoma samples were excluded from each respective study. Similarly, for comparisons in which normal versus carcinoma were made the adenoma samples were excluded from each respective study. The data was split between feces and tissue samples. Within the tissue groups the data was further divided between samples from the same individual (matched) and those from different individuals (unmatched). Where applicable for each study, predictions for adenoma and carcinoma were then tested for feces, matched tissue, and unmatched tissue.

***Obtaining Genera Relative Abundance and Selected Models:***
For the genera analysis of the RR, OTUs were added together based on the genus or lowest available taxonomic classification level and the total average counts, for 100 different subsamplings, were collected. The OTU based Random Forest Models using selected OTUS utilized a similar approach except that the OTUs were not aggregated together by taxonomic identity but kept as separate OTUs. OTU Random Forest models using the full community included all OTUs while those for the selected model included only those OTUs that had a taxonomic classification to a variable in the top 5 increased of top 5 decreased RR based on P-value.

### 2018-01-10 submission/manuscript.tex
update methods section to include top 5 component in workflow

\textbf{\emph{Study Analysis Overview:}} \(\alpha\)-diversity was first
assessed for differences between controls, adenoma, and carcinoma using
both linear mixed-effect models and RR. \(\beta\)-diversity was then
assessed for each individual study for differences between
control-adenoma and control-carcinoma. Next, four specific
CRC-associated genera (\emph{Fusobacterium}, \emph{Parvimonas},
\emph{Peptostreptococcus}, and \emph{Porphyromonas}) were assessed for
differences in RR for adenoma and carcinoma. We then built Random Forest
models based on all or CRC-associated genera and these models were
trained on one study then tested on the remaining studies, for every
study. A similar approach was then applied at the OTU level with the
exception that a 10-fold CV over 100 different models, based on random
80/20 splitting of the data, was used to generate a range of expected
AUCs. For these OTU based models, the CRC-associated genera included all
OTUs that had a taxonomic classification to \emph{Fusobacterium},
\emph{Parvimonas}, \emph{Peptostreptococcus}, or \emph{Porphyromonas}.
Finally, the power of each study was assessed for an effect size ranging
from 1\% to 30\% and an estimated sample size, for these effect sizes,
was generated based on 80\% power. For comparisons in which normal
versus adenoma were made the carcinoma samples were excluded from each
respective study. Similarly, for comparisons in which normal versus
carcinoma were made the adenoma samples were excluded from each
respective study. The data was split between feces and tissue samples.
Within the tissue groups the data was further divided between samples
from the same individual (matched) and those from different individuals
(unmatched). Where applicable for each study, predictions for adenoma
and carcinoma were then tested for feces, matched tissue, and unmatched
tissue.

\textbf{\emph{Obtaining CRC-associated Genera:}} For the CRC-associated
genera analysis of the RR, the total average counts, for 100 different
subsamplings, were collected for each respective OTU that had a genus
level taxonomic classification to \emph{Fusobacterium},
\emph{Parvimonas}, \emph{Peptostreptococcus}, and \emph{Porphyromonas}.
The OTU based Random Forest Models using CRC-associated genera utilized
a similar approach except that the OTUs were not aggregated together by
genus but kept as separate OTUs. OTU Random Forest models using the full
community included all OTUs while those using CRC-associated genera
included only those OTUs that had a genus level taxonomic classification
to \emph{Fusobacterium}, \emph{Parvimonas}, \emph{Peptostreptococcus},
and \emph{Porphyromonas}.

>>>>

\textbf{\emph{Study Analysis Overview:}} \(\alpha\)-diversity was first
assessed for differences between controls, adenoma, and carcinoma using
both linear mixed-effect models and RR. \(\beta\)-diversity was then
assessed for each individual study for differences between
control-adenoma and control-carcinoma. Next, all common genera were
assessed for differences in RR for adenoma and carcinoma. We then built
Random Forest models based on all or a selected model, based on the top
5 increased and top 5 decreased RR based on P-value, and these models
were trained on one study then tested on the remaining studies, for
every study. A similar approach was then applied at the OTU level with
the exception that a 10-fold CV over 100 different models, based on
random 80/20 splitting of the data, was used to generate a range of
expected AUCs. For these OTU-based models, the selected model included
all OTUs that had a taxonomic classification to a variable in the top 5
increased and top 5 decreased RR based on P-value. Finally, the power of
each study was assessed for an effect size ranging from 1\% to 30\% and
an estimated sample size, for these effect sizes, was generated based on
80\% power. For comparisons in which normal versus adenoma were made the
carcinoma samples were excluded from each respective study. Similarly,
for comparisons in which normal versus carcinoma were made the adenoma
samples were excluded from each respective study. The data was split
between feces and tissue samples. Within the tissue groups the data was
further divided between samples from the same individual (matched) and
those from different individuals (unmatched). Where applicable for each
study, predictions for adenoma and carcinoma were then tested for feces,
matched tissue, and unmatched tissue.

\textbf{\emph{Obtaining Genera Relative Abundance and Selected Models:}}
For the genera analysis of the RR, OTUs were added together based on the
genus or lowest available taxonomic classification level and the total
average counts, for 100 different subsamplings, were collected. The OTU
based Random Forest Models using selected OTUS utilized a similar
approach except that the OTUs were not aggregated together by taxonomic
identity but kept as separate OTUs. OTU Random Forest models using the
full community included all OTUs while those for the selected model
included only those OTUs that had a taxonomic classification to a
variable in the top 5 increased of top 5 decreased RR based on P-value.
